---
title: "beautier demo"
author: "Richèl J.C. Bilderbeek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{beautier demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

The purpose of `beautier` is to create a valid `BEAST2` XML input file
from its function argument. In this way, a scientific pipeline using 
`BEAST2` can be fully scripted, instead of using `BEAUti`'s GUI.

# Demonstration

First, `beautier` is loaded:

```{r}
library(beautier)
```

A BEAST2 XML input file needs an alignment (as BEAST2 infers
phylogenies and parameters on DNA sequences). This demonstration uses
a testing FASTA file used by `beautier`:

```{r}
fasta_filename <- get_beautier_path("test_output_0.fas")
```

We can display the alignment in the file:


```{r}
image(ape::read.FASTA(fasta_filename))
```

Now we can create our XML file:

```{r}
# The name of the file you intend to let BEAST2 run
output_filename <- tempfile(pattern = "beeast2", fileext = ".xml")

create_beast2_input_file(
  fasta_filename,
  output_filename
)
```

The file indeed is a BEAST2 input file:

```{r}
cat(readLines(output_filename), quote = FALSE, sep = '\n')
```

This XML input file can be read by BEAST2.
