---
title: "beautier demo"
author: "Richel J.C. Bilderbeek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{beautier demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

The purpose of `beautier` is to create a valid `BEAST2` XML input file
from its function argument. In this way, a scientific pipeline using 
`BEAST2` can be fully scripted, instead of using `BEAUti`'s GUI.

# Demonstration

First, `beautier` is loaded:

```{r}
library(beautier)
```

A BEAST2 XML input file needs an alignment (as BEAST2 infers
phylogenies and parameters on DNA sequences). in this demonstration,
a FASTA file with a random alignment is created:

```{r}
fasta_filename <- tempfile(pattern = "demo", fileext = ".fas")

beautier::create_random_fasta(
  n_taxa = 5, # species
  sequence_length = 20, # DNA nucleotides
  filename = fasta_filename
)
```

We can display the alignment in the file:


```{r}
image(ape::read.FASTA(fasta_filename))
```

Now we can create our XML file:

```{r}
# The name of the file you intend to let BEAST2 run
output_xml_filename <- tempfile(pattern = "demo", fileext = ".xml")

beautier::create_beast2_input_file(
  fasta_filename,
  output_xml_filename,
  tree_priors = create_yule_tree_prior(birth_rate_distribution = create_uniform_distr(id = 1))
)
```

(note: the line `tree_priors = create_yule_tree_prior(birth_rate_distribution = create_uniform_distr(id = 1))` will be removed somewhere close in the future)

The file indeed is a BEAST2 input file:

```{r}
print(readLines(output_xml_filename))
```

This XML input file can be read by BEAST2.
