---
title: "Examples"
author: "Richèl J.C. Bilderbeek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# Create files
for (filename in c("my_fasta.fas", "my_alignment.fas")) {
  file.copy(from = beautier::get_beautier_path("anthus_aco_sub.fas"), to = filename)
}

# Copy to local
for (filename in c("test_output_0.fas", "anthus_aco.fas", "anthus_nd2.fas")) {
  file.copy(from = beautier::get_beautier_path(filename), to = filename)
}
```

# Examples

For all examples, do load `beautier`:

```{r}
library(beautier)
```

All examples read the alignment from a FASTA file (usually `my_fasta.fas`) 
and create a BEAST2 input file called `my_beast.xml`.

## Example #1: all default

Using all default settings, only specify a DNA alignment.

```{r}
create_beast2_input_file(
  "test_output_0.fas",
  "my_beast.xml"
)
```

All other parameters are set to their defaults, as in BEAUti.

## Example #3: JC69 site model

```{r}
create_beast2_input_file(
  "my_alignment.fas",
  "my_beast.xml",
  site_model = create_jc69_site_model()
)
```

## Example #4: Relaxed clock log normal

```{r example_4}
create_beast2_input_file(
  "my_alignment.fas",
  "my_beast.xml",
  clock_model = create_rln_clock_model()
)
```

## Example #5: Birth-Death tree prior

```{r example_5}
create_beast2_input_file(
  "my_alignment.fas",
  "my_beast.xml",
  tree_prior = create_bd_tree_prior() 
)
```

## Example #6: Yule tree prior with a normally distributed birth rate

```{r example_6}
create_beast2_input_file(
  "my_alignment.fas",
  "my_beast.xml",
  tree_prior = create_yule_tree_prior(
    birth_rate_distr = create_normal_distr()
  ) 
)
```

Thanks to Yacine Ben Chehida for this use case

## Example #7: HKY site model with a non-zero proportion of invariants

```{r example_7}
create_beast2_input_file(
  "my_alignment.fas",
  "my_beast.xml",
  site_model = create_hky_site_model(
    gamma_site_model = create_gamma_site_model(prop_invariant = 0.5)
  )
)
```

Thanks to Yacine Ben Chehida for this use case

## Example #8: Strict clock with a known clock rate

```{r example_8}
create_beast2_input_file(
  "my_alignment.fas",
  "my_beast.xml",
  clock_model = create_strict_clock_model(
    clock_rate_param = 0.5
  ) 
)
```

Thanks to Paul van Els and Yacine Ben Chehida for this use case.

## Example #9: Use MRCA prior

Since `v1.13` it is supported to specify 
an MRCA ('Most Recent Common Ancestor') prior.

```{r example_9}
create_beast2_input_file(
  "my_alignment.fas",
  "my_beast.xml",
  mrca_prior = create_mrca_prior()
)
```

```{r example_9_fixed_crown_age}
create_beast2_input_file(
  "my_alignment.fas",
  "my_beast.xml",
  mrca_prior = create_mrca_prior(
    mrca_distr = create_normal_distr(
      mean = 10.0,
      sigma = 0.01
    )
  )
)
```

## Example #10: Use MRCA prior to specify a close-to-fixed crown age

Since `v1.13` it is supported to specify 
an MRCA ('Most Recent Common Ancestor') prior.

With an MRCA  prior, it is possible
to specify a close-to-fixed crown age:

```{r example_10}
crown_age <- 15
create_beast2_input_file(
  "my_alignment.fas",
  "my_beast.xml",
  mrca_prior = create_mrca_prior(
    is_monophyletic = TRUE,
    mrca_distr = create_normal_distr(
      mean = crown_age,
      sigma = 0.001
    )
  )
)
```

## Example #11: Two alignments

```{r example_11}
# Deprecated support for now
if (1 == 2) {
  create_beast2_input_file(
    c("anthus_aco.fas", "anthus_nd2.fas"),
    "my_beast.xml"
  )
}
```

Thanks to Paul van Els for this use case and supplying these FASTA files.

## Example #12: Two alignments, different site models

```
# Deprecated support for now
beautier::create_beast2_input_file(
  c("anthus_aco.fas", "anthus_nd2.fas"),
  "my_beast.xml",
  site_models = list(
    create_hky_site_model(), 
    create_tn93_site_model()
  )
)
```

Since `v1.12` this it is supported to have two alignments with different site models, clock models and tree priors.

Thanks to Paul van Els for this use case.

## Example #13: Two alignments, shared clock model

```{r example_13}
# Deprecated support for now
if (1 == 2) {
  beautier::create_beast2_input_file(
    c("anthus_aco.fas", "anthus_nd2.fas"),
    "my_beast.xml",
    clock_models = list(
      create_strict_clock_model(id = "anthus_aco"), 
      create_strict_clock_model(id = "anthus_aco")
    )
  )
}
```

In a future version, it will be supported to have two alignments with shared site models, clock models and tree priors.

Thanks to Yacine Ben Chehida for this use case.

```{r cleanup, include = FALSE}
# Cleaning up
filenames <- c(
  "my_fasta.fas", 
  "my_alignment.fas",
  "test_output_0.fas", 
  "anthus_aco.fas", 
  "anthus_nd2.fas",
  "my_beast.xml"
)
for (filename in filenames) {
  if (file.exists(filename)) file.remove(filename)
}
```
